Modeling Intervention Strategies for United States TB Control

### 2013-10-14 in_progress/presentations/paper/draft.tex
minor edit
changed intervention percentages to 5, 10, 25, and 50 percent

\subsection{Additional Tracking Capabilities}
In order to refine the tracking capabilities of the
Hill Model, the original differential equations used to describe TB spread were
separated into their component parts and each section was tracked separately.
These components were made into compartments, tracked by differential equations
detailed in the appendix. These equations allowed the model to track the total
number of TB cases, TB deaths, and  natural deaths. Further, the model tracks,
the sourced cost on the US HCS. Progressions into
active TB due to activations of LTBI or exogenous infection were tracked,
allowing for sourced incidence data to be generated. The entering cases of LTBI
(acute or chronic) were also tracked. In the case of intervention testing, the
number of cured cases of entering LTBI and the number of untreated entering LTBI
were both tracked. Cured cases of LTBI entering, TB deaths, total TB cases, and
total cost were also tracked discounted at 3\% annually. This discounting was
converted to a continuous differential equation for use within the model. In
general, incidence data is calculated in the same way in the extended hill as it
was in the basic Hill.

@@

\subsection{Intervention Analysis}
The primary interventions analyzed by the extended Hill were those that analyzed
curing various percentages of entering LTBI cases. Four indicative percentages
chosen were $20\%$, $35\%$, $50\%$, and $65\%$. Note that the Hill model does
not distinguish documented immigration from undocumented immigration, and as
such estimates of entering LTBI cure rates higher than $65\%$ become much more
difficult to achieve. It was seen that in this model as well as in the basic
Hill, no analyzed intervention predicted elimination by 2100. In order to obtain
elimination by 2100, at least 95\% of entering LTBI cases had to be cured, which
is practically impossible. However, it was seen that curing entering cases of
LTBI resulted in a net US HCS cost per case averted of \$67,722.96 at 2025, \$28,759.00 at 2050,
and \$17,967.39 at 2100 (assuming 20\% reduction with one cure costing \$XXXX; see Table
\ref{tab:cpcaRes} for other percentages). Given the variable nature of LTBI
treatment cost, the model code is extendible such that a user can adjust these
costs themselves to explore more specific methods of curing entering LTBI.
Further, it was also found that the relationship between total incidence at 2100
and percentage of incoming LTBI cases cured was linear, and from this estimates
were made of the yearly average US HCS savings garnered by curing one case of
entering LTBI over the time scale 2000 to 2025, 2000 to 2050, and 2000 to 2100.
This value peaked at \$1.026 million at 2100 (20\% reduction). This illustrates
that it would be cost saving to cure cases of LTBI at the cost of \$1.026
million ``2000'' dollars over the time period 2000-2100. These intervention
strategies also resulted in 23,870; 41,952; and 60,189 fewer cases of TB seen in
the US, and 2,056; 3,614; and 5,185 fewer TB deaths, for 20\%, 35\%, and 50\%
reduction, respectively.

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Sat Aug 24 18:08:52 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red20 & red35 & red50 & red65 \\ 
  \hline
2000 & ND & ND & ND & ND \\ 
  2025 & 67722.96 & 67527.15 & 67332.53 & 67138.90 \\ 
  2050 & 28759.00 & 28581.09 & 28404.13 & 28227.50 \\ 
  2075 & 20876.80 & 20709.13 & 20541.77 & 20374.49 \\ 
  2100 & 17967.39 & 17804.94 & 17642.64 & 17480.33 \\ 
   \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by 20\%, 35\%, 50\%, and 65\% (in dollar per case)} 
\label{tab:cpcaRes}
\end{table}
Several other intervention strategies analyzed by the Hill Model were refined
with the Extended Hill, and economic properties about each of them were tracked.
The results for these interventions, which are less effective than curing
entering LTBI cases across the board, are given in the appendix. 
\subsection{Agent Based Evaluation}
The Agent Based Model allowed the statistical properties of the system to be
analyzed and verified. In particular, they illustrated that the deterministic
Hill model provides a robust and consistent statistical measure of TB epidemic
behavior in the US conditions. We found that the distribution of incidence and
final population sizes were normal, with mean accurate to the deterministic
model and standard deviations given in table X, in the appendix.

@@

\subsection{Statistical Qualities of the Agent Based Model}
\subsection{Finer Intervention Analysis}
\subsection{Latin Hypercube Sampling}
The model parameters and variable names referred to in this analysis are listed below, along with best-fit or estimated values.  Following the example of the Hill model, we generated a Latin Hypercube Sample varying 18 of the input parameters.  From these parameters, 16 are identical to the input parameters varied in the sensitivity analysis of the original Hill model, and the remaining two parameters, $C_{A}$ and $C_{L}$, are variables for the average cost of Active and Latent TB treatment in the US.   Probability distributions for the original 16 parameters of the Hill model were all set to be Triangular, with mode at the best fit value and end points at the 2.5 and 97.5 percentile values reported in the Hill model.  Probability distributions for $C_{A}$ and $C_{L}$ were set to be Uniform, with range +/- 10\% of the estimated value.  All probabilty distributions used to generate the Latin Hypercube are listed in Table 1.\\

>>>>

\subsection{Additional Tracking Capabilities}
In order to refine the tracking capabilities of the
Hill Model, the original differential equations used to describe TB spread were
separated into their component parts and each section was tracked separately.
These components were made into compartments, tracked by differential equations
detailed in the appendix. These equations allowed the model to track the total
number of TB cases, TB deaths, and  natural deaths. Further, the model tracks,
the sourced cost on the US HCS. Progressions into
active TB due to activations of LTBI or exogenous infection were tracked,
allowing for sourced incidence data to be generated. The entering cases of LTBI
(acute or chronic) were also tracked. In the case of intervention testing, the
number of cured and untreated cases of entering LTBI were both tracked. Cured
cases of LTBI entering, TB deaths, total TB cases, and total cost were also
tracked discounted at 3\% annually. This discounting was converted to a
continuous differential equation for use within the model. In general, incidence
data is calculated in the same way in the extended hill as it was in the basic
Hill.

@@

\subsection{Intervention Analysis}
The primary interventions analyzed by the extended Hill were those that analyzed
curing various percentages of entering LTBI cases. Four indicative percentages
chosen were $5\%$, $10\%$, $25\%$, and $50\%$. Note that the Hill model does
not distinguish documented immigration from undocumented immigration, and as
such estimates of entering LTBI cure rates higher than $50\%$ become much more
difficult to achieve. It was seen that in this model as well as in the basic
Hill, no analyzed intervention predicted elimination by 2100. In order to obtain
elimination by 2100, at least 95\% of entering LTBI cases had to be cured, which
is practically impossible. However, it was seen that curing entering cases of
LTBI resulted in a net US HCS cost per case averted of \$67,654.07 at 2025, \$28,699.15 at 2050,
and \$17,912.95 at 2100 (assuming 25\% reduction with each cure costing \$800; see Table
\ref{tab:cpcaArt} for other percentages). Given the variable nature of LTBI
treatment cost, the model code is extendible such that a user can adjust these
costs themselves to explore more specific methods of curing entering LTBI.
Further, it was also found that the relationship between total incidence at 2100
and percentage of incoming LTBI cases cured was linear, and from this estimates
were made of the yearly average US HCS savings garnered by curing one case of
entering LTBI over the time scale 2000 to 2025, 2000 to 2050, and 2000 to 2100.
This value peaked at \$1.283 billion at 2100 (25\% reduction). This illustrates
that it would be cost saving to cure cases of LTBI at the cost of \$1.283
billion ``2000'' dollars over the time period 2000-2100. These intervention
strategies also resulted in 11,900; 29,880; and 60,189 fewer cases of TB seen in
the US, and 1,025; 2,573; and 5,185 fewer TB deaths, for 10\%, 25\%, and 50\%
reduction, respectively.

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red5 & red10 & red25 & red50 \\ 
  \hline
2000 & ND & ND & ND & ND \\ 
  2025 & 67908.81 & 67855.46 & 67654.07 & 67332.53 \\ 
  2050 & 28938.23 & 28878.70 & 28699.15 & 28404.13 \\ 
  2075 & 21044.88 & 20989.16 & 20820.53 & 20541.77 \\ 
  2100 & 18130.11 & 18076.11 & 17912.95 & 17642.64 \\ 
   \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by X percent (in dollar per case)} 
\label{tab:cpcaArt}
\end{table}

Several other intervention strategies analyzed by the Hill Model were refined
with the Extended Hill, and economic properties about each of them were tracked.
The results for these interventions, which are less effective than curing
entering LTBI cases across the board, are given in the appendix. 
\subsection{Agent Based Evaluation}
The Agent Based Model allowed the statistical properties of the system to be
analyzed and verified. In particular, they illustrated that the deterministic
Hill model provides a robust and consistent statistical measure of TB epidemic
behavior in the US conditions. We found that the distribution of incidence and
final population sizes were normal, with mean accurate to the deterministic
model and standard deviations given in table X, in the appendix.

@@

\subsection{Statistical Qualities of the Agent Based Model}
\subsection{Finer Intervention Analysis}

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red5 & red10 & red25 & red50 \\ 
  \hline
2000 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  2025 & 1115.12 & 2231.70 & 5592.43 & 11227.57 \\ 
  2050 & 3406.41 & 6821.27 & 17117.18 & 34447.71 \\ 
  2075 & 4976.77 & 9967.15 & 25021.33 & 50389.58 \\ 
  2100 & 5942.02 & 11900.90 & 29880.03 & 60189.49 \\ 
   \hline
\end{tabular}
\caption{Cases of TB Averted by Reducing Incoming LTBI by X percent} 
\label{tab:caRed}
\end{table}

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red5 & red10 & red25 & red50 \\ 
  \hline
2000 & ND & ND & ND & ND \\ 
  2025 & 67908.81 & 67855.46 & 67654.07 & 67332.53 \\ 
  2050 & 28938.23 & 28878.70 & 28699.15 & 28404.13 \\ 
  2075 & 21044.88 & 20989.16 & 20820.53 & 20541.77 \\ 
  2100 & 18130.11 & 18076.11 & 17912.95 & 17642.64 \\ 
   \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by X percent (in dollar per case)} 
\label{tab:cpcaRed}
\end{table}

\subsection{Latin Hypercube Sampling}
The model parameters and variable names referred to in this analysis are listed below, along with best-fit or estimated values.  Following the example of the Hill model, we generated a Latin Hypercube Sample varying 18 of the input parameters.  From these parameters, 16 are identical to the input parameters varied in the sensitivity analysis of the original Hill model, and the remaining two parameters, $C_{A}$ and $C_{L}$, are variables for the average cost of Active and Latent TB treatment in the US.   Probability distributions for the original 16 parameters of the Hill model were all set to be Triangular, with mode at the best fit value and end points at the 2.5 and 97.5 percentile values reported in the Hill model.  Probability distributions for $C_{A}$ and $C_{L}$ were set to be Uniform, with range +/- 10\% of the estimated value.  All probabilty distributions used to generate the Latin Hypercube are listed in Table 1.\\
